---
title: "Profile Reconstruction"
author: "Thanina Chabane"
date: "2024-06-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Expression profiles reconstruction

```{r}
source("profile_reconstruction_source.R")
```

## Condition CTRL

```{r}
sample_1340 <- readRDS("data/clean/samp_1340_bio_ordered.Rds")
```

```{r}
gene = "Mup11"
```

```{r}

table <- RunLinearRegionning(sample_1340, 40, start_region = "1", target_region = "6", GOI = gene, seed =20, return_df = T)

ggplot(table, aes(x = scaled_dist, y = expression, fill = GOI)) +
    geom_smooth(method = "loess", colour="black") +
    theme_minimal() + ggtitle(label = samp1$GOI) + 
    xlab("Minimal distance between portal and central") + 
    theme(plot.title = element_text(hjust = 0.5, size=16, face="bold")) 
```

Beware, this function may return an error. if so, do run each command individually (plots will be generated regardless of the error)

```{r}
 RunHexagonalRegionning(sample_1340, 20, left_region = "1", right_region = "6", GOI = "Spp1", distance = 400)
  RunHexagonalRegionning(sample_1340, 20, left_region = "1", right_region = "6", GOI = "Cyp2f2", distance = 400)
  RunHexagonalRegionning(sample_1340, 20, left_region = "1", right_region = "6", GOI = "Mup11", distance = 400)
  RunHexagonalRegionning(sample_1340, 20, left_region = "1", right_region = "6", GOI = "Glul", distance = 400)
```

## Condition STIM


```{r}
sample_1352 <- readRDS("data/clean/samp_1352_bio_ordered.Rds")
```

```{r}
gene = "Mup11"
```

```{r}

table <- RunLinearRegionning(sample_1352, 40, start_region = "1", target_region = "6", GOI = gene, seed =20, return_df = T)

ggplot(table, aes(x = scaled_dist, y = expression, fill = GOI)) +
    geom_smooth(method = "loess", colour="black") +
    theme_minimal() + ggtitle(label = samp1$GOI) + 
    xlab("Minimal distance between portal and central") + 
    theme(plot.title = element_text(hjust = 0.5, size=16, face="bold")) 
```

Beware, this function may return an error. if so, do run each command individually (plots will be generated regardless of the error)

```{r}
 RunHexagonalRegionning(sample_1352, 20, left_region = "1", right_region = "6", GOI = "Spp1", distance = 400)
  RunHexagonalRegionning(sample_1352, 20, left_region = "1", right_region = "6", GOI = "Cyp2f2", distance = 400)
  RunHexagonalRegionning(sample_1352, 20, left_region = "1", right_region = "6", GOI = "Mup11", distance = 400)
  RunHexagonalRegionning(sample_1352, 20, left_region = "1", right_region = "6", GOI = "Glul", distance = 400)
```


## Condition DaKO


```{r}
sample_1353 <- readRDS("data/clean/samp_1353_bio_ordered.Rds")
```

```{r}
gene = "Mup11"
```

```{r}

table <- RunLinearRegionning(sample_1353, 40, start_region = "1", target_region = "6", GOI = gene, seed =20, return_df = T)

ggplot(table, aes(x = scaled_dist, y = expression, fill = GOI)) +
    geom_smooth(method = "loess", colour="black") +
    theme_minimal() + ggtitle(label = samp1$GOI) + 
    xlab("Minimal distance between portal and central") + 
    theme(plot.title = element_text(hjust = 0.5, size=16, face="bold")) 
```

Beware, this function may return an error. if so, do run each command individually (plots will be generated regardless of the error)

```{r}
 RunHexagonalRegionning(sample_1353, 20, left_region = "1", right_region = "6", GOI = "Spp1", distance = 400)
  RunHexagonalRegionning(sample_1353, 20, left_region = "1", right_region = "6", GOI = "Cyp2f2", distance = 400)
  RunHexagonalRegionning(sample_1353, 20, left_region = "1", right_region = "6", GOI = "Mup11", distance = 400)
  RunHexagonalRegionning(sample_1353, 20, left_region = "1", right_region = "6", GOI = "Glul", distance = 400)
```

## Condition DaKO STIM


```{r}
sample_1354 <- readRDS("data/clean/samp_1354_bio_ordered.Rds")
```

```{r}
gene = "Mup11"
```

```{r}

table <- RunLinearRegionning(sample_1354, 40, start_region = "1", target_region = "6", GOI = gene, seed =20, return_df = T)

ggplot(table, aes(x = scaled_dist, y = expression, fill = GOI)) +
    geom_smooth(method = "loess", colour="black") +
    theme_minimal() + ggtitle(label = samp1$GOI) + 
    xlab("Minimal distance between portal and central") + 
    theme(plot.title = element_text(hjust = 0.5, size=16, face="bold")) 
```

Beware, this function may return an error. if so, do run each command individually (plots will be generated regardless of the error)

```{r}
 RunHexagonalRegionning(sample_1354, 20, left_region = "1", right_region = "6", GOI = "Spp1", distance = 400)
  RunHexagonalRegionning(sample_1354, 20, left_region = "1", right_region = "6", GOI = "Cyp2f2", distance = 400)
  RunHexagonalRegionning(sample_1354, 20, left_region = "1", right_region = "6", GOI = "Mup11", distance = 400)
  RunHexagonalRegionning(sample_1354, 20, left_region = "1", right_region = "6", GOI = "Glul", distance = 400)
```